SAT Solving for Vesicle Traffic Systems in Cells

### 2017-07-15 paper/bio.tex
added srivas wworkline
fixed table refereing
V6
fixed table and added line numb and fixed description table 1

\textbf{The cell as a transport graph:} We consider a cell to be a
collection of compartments (nodes) and vesicles (edges), thus defining
a transport graph.
%
Compartments represent the large membrane-bound organelles, which are
connected by constant fluxes of small transport vesicles.
%
Every compartment or vesicle has a set of molecules associated with
it.
%
These include SNAREs and other biomolecules that regulate membrane
traffic, or that play the role of passive cargo.
%
Here we consider an abstract set of molecules and do
not explicitly associate them with particular protein varieties.

\textbf{Molecular flows and steady state:} The total amount of each
molecular type on each compartment can increase or decrease: this is
due to gain or loss of that molecular type driven by incoming or
outgoing vesicles.
%
Each edge is thus associated with a flux of all the molecular types
carried by the corresponding vesicle.
%
We assume the cell is in a steady state where each compartments
composition does not vary over time, meaning that all incoming and
outgoing fluxes are balanced for each molecular type at each
compartment.

\textbf{Vesicle targeting driven by molecular interactions:} The model
specification includes a description of how molecular properties
influence vesicle transport.
%
A vesicle can only contain a set of molecules that are present
on its source compartment.
%
Once a vesicle has budded out of the source, the molecules it carries
determine its properties.
%
In particular, for any given pair of a vesicle and a compartment, the
set of molecules that label the former and latter determine whether
the vesicle will fuse to that compartment.
%
Biophysically, fusion requires a direct physical interaction between
at least one molecular type on the vesicle and one molecular type on
the compartment.
%
The list of molecular pairs that can drive a fusion event is given an
a fusion pairing matrix.

@@

\textbf{Synthesis:} We now combine all the above ingredients.
%
Given a particular
transport graph, a particular labeling of all the compartments and edges,
a particular fusion pairing matrix, and a particular regulatory model we do
the following.
\begin{enumerate}
\item We determine which molecules are ``active" on every
compartment or vesicle.
\item or every vesicle fusing to a compartment, we
determine whether there exists an active pair (one molecule on the vesicle,
one on the compartment) which drives that fusion event.
\item For every vesicle-compartment pair where the vesicle does not fuse to the
compartment, we verify that there is no pairing of active molecules on the
vesicle and compartment that could drive their fusion.
\item We verify that every molecular type entering a compartment also leaves the compartment,
and also that every molecular type entering a set of compartments also
leaves that set; this is the steady state condition.
%
This tells us that the
particular graph and molecular labeling does represent an allowed steady
state configuration of a VTS.
\end{enumerate}

>>>>

\textbf{The cell as a transport graph:} We consider a cell to be a
collection of compartments (nodes) and vesicles (edges), thus defining
a transport graph.
%
Compartments represent the large membrane-bound organelles, which are
connected by constant fluxes of small transport vesicles.
%
Every compartment or vesicle has a set of molecules associated with
it.
%
These include SNAREs and other biomolecules that regulate membrane
traffic, or that play the role of passive cargo.
%
Here we consider an abstract set of molecules and do
not explicitly associate them with particular protein varieties.\textbf{But consider them to be either of  
the two categories Q and R.}
\textbf{Molecular flows and steady state:} The total amount of each
molecular type on each compartment can increase or decrease: this is
due to gain or loss of that molecular type driven by incoming or
outgoing vesicles.
%
Each edge is thus associated with a flux of all the molecular types
carried by the corresponding vesicle.
%
We assume the cell is in a steady state where each compartments
composition does not vary over time, meaning that all incoming and
outgoing fluxes are balanced for each molecular type at each
compartment.

\textbf{Vesicle targeting driven by molecular interactions:} The model
specification includes a description of how molecular properties
influence vesicle transport.
%
A vesicle can only contain a set of molecules that are present
on its source compartment.
%
Once a vesicle has budded out of the source, the molecules it carries
determine its properties.
%
In particular, for any given pair of a vesicle and a compartment, the
set of molecules that label the former and latter determine whether
the vesicle will fuse to that compartment.
%
Biophysically, fusion requires a direct physical interaction between
at least one molecular type on the vesicle and one molecular type on
the compartment.
%
The list of molecular pairs that can drive a fusion event is given an
a fusion pairing matrix. \textbf{For convenience, we divide Q and R SNAREs evenly 
and allow only opposite types to become a pair.}

@@

\textbf{Synthesis:} We now combine all the above ingredients.
%
Given a particular
transport graph, a particular labeling of all the compartments and edges,
a particular fusion pairing matrix, and a particular regulatory model we do
the following.
\begin{enumerate}
\item We determine which molecules are ``active" on every
compartment or vesicle.
\item or every vesicle fusing to a compartment, we
determine whether there exists an active pair (one molecule on the vesicle,
one on the compartment) which drives that fusion event.
\item For every vesicle-compartment pair where the vesicle does not fuse to the
compartment, we verify that there is no pairing of active molecules on the
vesicle and compartment that could drive their fusion.
\item We verify that every molecular type entering a compartment also leaves the compartment,
and also that every molecular type entering a set of compartments also
leaves that set; this is the steady state condition.
%
This tells us that the
particular graph and molecular labeling does represent an allowed steady
state configuration of a VTS.
\end{enumerate}
%The fusion is a process where SNAREs on the vesicles target compartment is driven by the zipping together of vesicle
%and compartment SNAREs [8, 9].
%The SNAREs can be 
%Typically a Q-SNARE triple-helix is confined to one mem-
%brane surface (e.g the compartment), while an R-SNARE single helix is confined to its fusion
%partner (e.g. the vesicle). When these two types of SNAREs are in close proximity they zip
%together into a four-helix bundle, forcing the vesicle to fuse with the compartment. There are
%many types of Q-SNAREs and R-SNAREs spread between different vesicles and compart-
%ments. Across eukaryotes, there are 20 broad SNARE varieties, though individual species can
%contain higher numbers (humans have 41) [12]. Only certain Q and R-SNARE combinations
%can zip into complexes, regulating precisely which vesicles fuse with which targets. For our
%analysis, we consider a Q-SNARE triple helix as a single complex and the R-SNARE as its cog-
%nate partner. We refer to zipping as SNARE pairing.

\begin{itemize}
\item \srivas{Detailed BIO to Graph problem}
\end{itemize}

### 2017-07-15 paper/encoding.tex
V6
fixed table and added line numb and fixed description table 1

\begin{align}
  \bigwedge\limits_{i,j,q,m} e_{i,j,q,m} \limplies (n_{i,m} \land n_{j,m} )\tag{V4}\label{eq:c0}
\end{align}
Self edges are not allowed. 
\begin{align}
   \bigwedge\limits_{i,q} \neg e_{i,i,q}\tag{V5}\label{eq:c2}
\end{align}
Since a molecule cannot be both SNARE and tether,
we set diagonal blocks in pairing matrix to be 0's.
It reduces the search space.
% Condition on $p_{kk'}$. \ashu{why???} .
\begin{align}
  \bigwedge\limits_{(x < M/2 \, \land  \, y < M/2) \lor  (x >= M/2 \land y >= M/2)} \neg p(x,y)
  \tag{V6}\label{eq:c3}
\end{align}

@@

\begin{enumerate}[label=\Alph*]
\item Every present molecule is considered to be active.
\item Activity of molecules on the nodes is based on Boolean function of presence of other molecules. 
\item Activity of molecules on the edges is based on Boolean function of presence of other molecules
\item Activity of molecules both on the edges and nodes is based on Boolean function of presence of other molecules.
\item Activity of molecules on the edges is driven by pairing inhibition.
\item Activity of molecules on the nodes is based on Boolean function of presence of other molecules and on edge by pairing inhibition.
%\item Version F with additional constraint that every edge is distinct.\ashu{encode this??}
%\item Activity of molecules on the nodes is based on Boolean function of presence of other molecules and on edge by pairing inhibition with every edge is distinct.
%\item Activity of molecules on the nodes is based on Boolean function of presence of other moleculess and on edge by pairing inhibition with self-edges are allowed.
%\item Version F but self-edges are allowed.
\end{enumerate}
%
The version (A) is most basic where every present molecule is active.
%
In versions (B-D,F),
most general restriction is activity of the molecule is dependent on
the other molecules presence on the node.
%
In the versions (E,F), activity of the molecules on the edges
is dictated by pairing matrix; we call this process ``pairing
inhibition".
%
In the column 2-3 of table~\ref{table1}, we list the constraints involved
in each version.

>>>>

\begin{align}
  \bigwedge\limits_{i,j,q,m} e_{i,j,q,m} \limplies (n_{i,m} \land n_{j,m} )\tag{V4}\label{eq:c0}
\end{align}
Self edges are not allowed. 
\begin{align}
   \bigwedge\limits_{i,q} \neg e_{i,i,q}\tag{V5}\label{eq:c2}
\end{align}
Since we have fixed first half as Q-SNAREs and rest as R-SNAREs,
we set diagonal blocks in pairing matrix to be 0's.
%It reduces the search space.
% Condition on $p_{kk'}$. \ashu{why???} .
\begin{align}
  \bigwedge\limits_{(x < M/2 \, \land  \, y < M/2) \lor  (x >= M/2 \land y >= M/2)} \neg p(x,y)
  \tag{V6}\label{eq:c3}
\end{align}

@@

\begin{enumerate}[label=\Alph*]
\item Every present molecule is considered to be active.
\item Activity of molecules on the nodes is based on Boolean function of presence of other molecules. 
\item Activity of molecules on the edges is based on Boolean function of presence of other molecules
\item Activity of molecules both on the edges and nodes is based on Boolean function of presence of other molecules.
\item Activity of molecules on the edges is driven by pairing inhibition.
\item Activity of molecules on the nodes is based on Boolean function of presence of other molecules and on edge by pairing inhibition.
%\item Version F with additional constraint that every edge is distinct.\ashu{encode this??}
%\item Activity of molecules on the nodes is based on Boolean function of presence of other molecules and on edge by pairing inhibition with every edge is distinct.
%\item Activity of molecules on the nodes is based on Boolean function of presence of other moleculess and on edge by pairing inhibition with self-edges are allowed.
%\item Version F but self-edges are allowed.
\end{enumerate}
%
For the version A, every present molecule is in active state. In the table~\ref{table1}, we list the constraints involved in each version. Constraints V1-V8, R1-R2 and D1-D4 are same for every version but the difference is the constraint deciding the activity of the molecules. One type of restriction is where the activity of the present molecule is dependent on the presence of the other molecules. For example in version B, E, F activity of the molecules on the node is a boolean function of the presence of other molecules on that node; Anb. Similarly for the case of the edge in version C and E; Aeb. In the case of versions F the activity of the molecules on the edges is described by pairing matrix, if all the pairs of the present molecule are present on the same edge than all of them inhibit each other resulting in making all the molecules inactive, we term this process ``pairing inhibition".
%

### 2017-07-15 paper/experiments.tex
fixed table and added line numb and fixed description table 1

A. & \multirow{5}{*}{
     \makecell{ \ref{eq:f0}--\ref{eq:fuse2},\\
     \ref{eq:reach1},\ref{eq:reach2},\\
     \ref{eq:drop1}--\ref{eq:drop4}}
     } 
& \ref{eq:ann},\ref{eq:aen} & No graph & No graph \\ \cline{3-5}
B. & & \ref{eq:anb},\ref{eq:aen} & No graph & No graph \\ \cline{3-5}
C. & & \ref{eq:ann},\ref{eq:aeb} & 3-connected & 3-connected \\  \cline{3-5}
D. & & \ref{eq:ann},\ref{eq:aep} & No graph & No graph \\ \cline{3-5}
E. & & \ref{eq:anb},\ref{eq:aeb} & 2-connected & 3-connected \\ \cline{3-5}
F. & & \ref{eq:anb},\ref{eq:aep} & 4-connected & 4-connected \\ \hline
%G. & & \ref{eq:anb},\ref{eq:aeb},\ref{eq:ced} & 3-connected & 3-connected \\ \hline

@@

% Dashed lines are used in the resultant visual network output to
% specify the dropped edges to make the graph disconnected, which gives
% information about the connectivity of the graph.
\ashu{discuss table~\ref{tab:grph}}

\input{tbl-qf-graph}

>>>>

A. & \multirow{5}{*}{
     \makecell{ \ref{eq:f0}--\ref{eq:fuse2},\\
     \ref{eq:reach1},\ref{eq:reach2},\\
     \ref{eq:drop1}--\ref{eq:drop4}}
     } 
& \ref{eq:ann},\ref{eq:aen} & No graph & No graph \\ \cline{3-5}
B. & & \ref{eq:anb},\ref{eq:aen} & No graph & No graph \\ \cline{3-5}
C. & & \ref{eq:ann},\ref{eq:aeb} & 3-connected & 3-connected \\  \cline{3-5}
D. & & \ref{eq:anb},\ref{eq:aeb} & 2-connected & 3-connected \\ \cline{3-5}
E. & & \ref{eq:ann},\ref{eq:aep} & No graph & No graph \\ \cline{3-5}
F. & & \ref{eq:anb},\ref{eq:aep} & 4-connected & 4-connected \\ \hline
%G. & & \ref{eq:anb},\ref{eq:aeb},\ref{eq:ced} & 3-connected & 3-connected \\ \hline

@@

% Dashed lines are used in the resultant visual network output to
% specify the dropped edges to make the graph disconnected, which gives
% information about the connectivity of the graph.
%\ashu{discuss table~\ref{tab:grph}}
table~\ref{tab:grph} shows the result of the different variations of the VTS. We haven't found any graph for the version A with constraints Ann and Aen. Replacing constraint on the node of Ann with Anb (version B) does not affect the outcome and hence there is no graph that is a valid VTS following these constraints. Interestingly if we allow every present molecule to stay active (Ann) but constraint the edge by a boolean function (Aeb) the resultant graph is a 3-connected graph. Constraining the activity of both the edge and the nodes with the boolean function (Anb Aeb) results in 2 connected graphs. Again there is no graph for the case where every molecule on the node is active but the activity on the edge is driven by pairing inhibition (Aep). But if we replace the constraint on the node to be a boolean function (Anb) with Aep we have found that 4 connected graphs satisfies all the constraints. 
\input{tbl-qf-graph}
